The ChEMBL database as Linked Open Data

### 2012-05-01 article.tex
Migrated to the BMC template, as we are now submitting to JChemInf

% PLEASE USE THIS FILE AS A TEMPLATE
% Check file iosart2c.tex for more examples
%
% Journal:
%   Journal of Ambient Intelligence and Smart Environments (jaise)
%   Web Intelligence and Agent Systems: An International Journal (wias)
%   Semantic Web: Interoperability, Usability, Applicability (SW)
% IOS Press
% Latex 2e

% options: jaise|wias|sw
% add. options: [seceqn,secfloat,secthm,crcready,onecolumn]

\documentclass[sw]{iosart2c}

%\documentclass[sw]{iosart2c}
%\documentclass[wias]{iosart2c}
%\documentclass[jaise]{iosart2c}

\usepackage[T1]{fontenc}
\usepackage{times}%
\usepackage{natbib}% for bibliography sorting/compressing
%\usepackage{amsmath}
%\usepackage{endnotes}
\usepackage{graphicx}
%\usepackage{tikz}
%\usetikzlibrary{snakes,arrows,shapes}

%%%%%%%%%%% Put your definitions here

% \bibliographystyle{plain} 
\bibliographystyle{unsrt}

%%%%%%%%%%% End of definitions

\pubyear{0000}
\volume{0}
\firstpage{1}
\lastpage{1}

\begin{document}

\newcommand{\url}[1]{#1}

\begin{frontmatter}

%\pretitle{}
\title{The ChEMBL database as Linked Open Data}
\runningtitle{ChEMBL-RDF}
%\subtitle{}

%\review{}{}{}

% For one author:
\author[A]{\fnms{Egon} \snm{Willighagen}\thanks{Corresponding author.\\ E-mail: egon.willighagen@maastrichtuniversity.nl.}}
\runningauthor{E.L. Willighagen et al.}
% Two or more authors (order to be determined):
\author[A]{\fnms{Andra} \snm{Waagmeester}}
\author[B]{\fnms{Ola} \snm{Spjuth}},
\author[C]{\fnms{Peter} \snm{Ansell}},
\author[D]{\fnms{Antony} \snm{Williams}},
\author[D]{\fnms{Valery} \snm{Tkachenko}},
\author[E]{\fnms{Janna} \snm{Hastings}},
\author[F]{\fnms{John} \snm{Overington}},
\author[F]{\fnms{Anna} \snm{Gaulton}},
\author[F]{\fnms{Mark} \snm{Davies}},
\author[G]{\fnms{Bin} \snm{Chen}},
\author[G]{\fnms{David} \snm{Wild}}

\address[A]{Department of Bioinformatics - BiGCaT, Maastricht University, P.O. Box 616, UNS50 Box 19, NL-6200 MD, Maastricht, The Netherlands}
\address[B]{Department of Pharmaceutical Biosciences, Uppsala University, PO Box 591, SE-751 24, Uppsala, Sweden}
\address[C]{University of Queensland, St Lucia, Qld 4072, Australia}
\address[D]{Royal Society of Chemistry, 904 Tamaras Circle, Wake Forest, NC 27587, U.S.A.}
\address[E]{Chemoinformatics and Metabolism, European Bioinformatics Institute, POSTAL CODE, Hinxton, United Kingdom}
\address[F]{EMBL-European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom}
\address[G]{School of Informatics and Computing, Indiana University, Bloomington, IN, U.S.A.}

\begin{abstract}
Making data available as Linked Data using RDF is advantageous because it promotes integration with other web resources.
RDF documents can natively link to related data, and others can link back using URIs. RDF makes the data machine-readable,
using extensible vocabularies for additional information. This combination provides a new, open, interface for scientists to use ChEMBL.
This paper describes the continued conversion of data from the ChEMBL database into RDF triples.
This updated version of ChEMBL-RDF now uses recently introduced ontologies, including CHEMINF and CiTO,
exposes more information from the database, and is now available as dereferencable, Linked Data.
To demonstrate the new features, we present new use cases showing the benefits of integration with
other web resources using semantic web technologies.
\end{abstract}

\begin{keyword}
% \sep 
Resource Description Framework, ChEMBL
\end{keyword}

\end{frontmatter}

@@

\section{Introduction}\label{s1}

@@

\section{Methods}\label{s2}

The ChEMBL version used in this paper is ChEMBL 13, which was released on 29 February 2012.
The SQL datadump provided by EMBL was inserted into a local MySQL database. A set of SQL queries 
were then executed to generate RDF files using a set of PHP scripts. These scripts are Open Source 
and are available from the source code hosting service, GitHub~\citep{ChEMBLRDFGitHub}. This process 
has been outlined in a best practices note by the W3C Health Care and Life Sciences XXX Working Group~\cite{Marshall2012}.

@@

\section{Results}\label{s3}

@@

In addition to the PubMed identifiers used to link from literature references to Bio2RDF,
ChEMBL provides DOIs, which we use to link out to the RDF provided by CrossRef~\citep{Bilder2011}:

@@

\section{Applications}

@@

\section{Discussion}

@@

\section{Acknowledgements}

@@

\bibliography{article}

%%%%%%%%%%% The bibliography starts:
%\begin{thebibliography}{9}
%
%\bibitem{r1}
%
%\bibitem{r2}
%
%\end{thebibliography}

\end{document}

>>>>

\documentclass[10pt]{bmc_article}

\usepackage{a4wide}  % Formatting web addresses  
\usepackage{url}  % Formatting web addresses  
\usepackage{ifthen}  % Conditional 
\usepackage{multicol}   %Columns
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\urlstyle{rm}
\usepackage[citestyle=numeric-comp]{biblatex}                                                                     
\setlength{\bibitemsep}{0cm}                                                                                                                
\renewbibmacro{in:}{%                                                                                                                       
  \ifentrytype{article}{}{%                                                                                                                 
  \printtext{\bibstring{in}\intitlepunct}}}                                                                                                 
\addbibresource{article.bib}

%\usepackage{amsmath}
%\usepackage{endnotes}
\usepackage{graphicx}
%\usepackage{tikz}

\newboolean{publ}

%Review style settings
\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}

%Publication style settings
%\newenvironment{bmcformat}{\fussy\setboolean{publ}{true}}{\fussy}

\begin{document}
\begin{bmcformat}

%\pretitle{}
\title{The ChEMBL database as Linked Open Data}

\author{
  Egon L Willighagen$^{1}$% // FIXED FIRST AUTHOR
  \email{Egon L Willighagen - egon.willighagen@maastrichtuniversity.nl}
\and
  Andra Waagmeester$^1$%
  \email{Andra Waagmeester - FIXME}
\and
  Ola Spjuth$^2$%
  \email{Ola Spjuth - FIXME}
\and
  Peter Ansell$^3$%
  \email{Peter Ansell - FIXME}
\and
  Antony Williams$^4$%
  \email{Antony Williams - FIXME}
\and
  Valery Tkachenko$^4$%
  \email{Valery Tkachenko - FIXME}
\and
  Janna Hastings$^5$%
  \email{Janna Hastings - FIXME}
\and
  John Overington$^6$%
  \email{John Overington - FIXME}
\and
  Anna Gaulton$^6$%
  \email{Anna Gaulton - FIXME}
\and
  Mark Davies$^6$%
  \email{Mark Davies - FIXME}
\and
  Bin Chen$^7$%
  \email{Bin Chen - FIXME}
\and
  David Wild$^7$%
  \email{David Wild - FIXME}
}

\address{
    \iid(1)Department of Bioinformatics - BiGCaT, Maastricht University, P.O. Box 616, UNS50 Box 19, NL-6200 MD, Maastricht, The Netherlands  \\
    \iid(2)Department of Pharmaceutical Biosciences, Uppsala University, PO Box 591, SE-751 24, Uppsala, Sweden \\
    \iid(3)University of Queensland, St Lucia, Qld 4072, Australia \\
    \iid(4)Royal Society of Chemistry, 904 Tamaras Circle, Wake Forest, NC 27587, U.S.A. \\
    \iid(5)Chemoinformatics and Metabolism, European Bioinformatics Institute, POSTAL CODE, Hinxton, United Kingdom \\
    \iid(6)EMBL-European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom \\ 
    \iid(7)School of Informatics and Computing, Indiana University, Bloomington, IN, U.S.A.
}

\maketitle

\begin{abstract}
\paragraph*{Background:}
Making data available as Linked Data using RDF is advantageous because it promotes integration with other web resources.
RDF documents can natively link to related data, and others can link back using URIs. RDF makes the data machine-readable,
using extensible vocabularies for additional information. This combination provides a new, open, interface for scientists to use ChEMBL.
\paragraph*{Results:}
This paper describes the continued conversion of data from the ChEMBL database into RDF triples.
This updated version of ChEMBL-RDF now uses recently introduced ontologies, including CHEMINF and CiTO,
exposes more information from the database, and is now available as dereferencable, Linked Data.
To demonstrate the new features, we present new use cases showing the benefits of integration with
other web resources using semantic web technologies.
\paragraph*{Conclusions:}
\end{abstract}

@@

\section*{Introduction}\label{s1}

@@

\section*{Methods}\label{s2}

The ChEMBL version used in this paper is ChEMBL 13, which was released on 29 February 2012.
The SQL datadump provided by EMBL was inserted into a local MySQL database. A set of SQL queries 
were then executed to generate RDF files using a set of PHP scripts. These scripts are Open Source 
and are available from the source code hosting service, GitHub~\cite{ChEMBLRDFGitHub}. This process 
has been outlined in a best practices note by the W3C Health Care and Life Sciences XXX Working Group~\cite{Marshall2012}.

@@

\section*{Results}\label{s3}

@@

In addition to the PubMed identifiers used to link from literature references to Bio2RDF,
ChEMBL provides DOIs, which we use to link out to the RDF provided by CrossRef~\cite{Bilder2011}:

@@

\section*{Applications}

@@

\section*{Discussion}

@@

\section*{Authors contributions}
EW initiated the project, created the initial RDF version of the ChEMBL data, and encouraged the use cases.
AW extended CitedIn to support citation info in ChEMBL-RDF.
OS, AW, and VT developed the nearest neighbor application of Bioclipse and ChemSpider.
JH supported the project with CHEMINF representations.
PA integrated ChEMBL-RDF into Bio2RDF.
JO, AG, and MD critically validated and summarized the ChEMBL data statistics.
All authors contributed to the continued development of ChEMBL-RDF, as well as to the writing of the paper,
and approved the final version.

\section*{Acknowledgements}

@@

{\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
\printbibliography
}

\ifthenelse{\boolean{publ}}{\end{multicols}}{}

\end{bmcformat}
\end{document}
